---
title: "Get Performance data of each student"
output: html_notebook
---

The “Rendimiento” databases contain information on the performance of each student in primary and secondary school, including data on grade point average and percentage of attendance.

################################################################################
### 1. Load data and packages

```{r, message=FALSE, warning=FALSE}
# Load packages
library(data.table)
library(tidyverse)
```

```{r, eval=FALSE}
# Load dropouts data
dropout_data <- readRDS("2_data/saved_R_objects/02_dropout_data_2009_2018.RDS")
```

**DO NOT RUN THIS**
```{r, eval=FALSE}
# Changes to Performance 2014 dataset
#-------------------------------------------------------------------------------
# create file path
csv_filepath <- paste0("2_data/data_rendimiento/rendimiento_2014_publ_original.csv.gz")
# read csv from compressed file
rendim_2014_original <- fread(csv_filepath, sep = ";", encoding = "UTF-8")
# changing variable name to match other years
names(rendim_2014_original)[31] <- "SIT_FIN_R"
# save base with changes
write_csv2(rendim_2014_original,"2_data/data_rendimiento/rendimiento_2014_publ.csv")
# removing object matsup_2019_origin
rm(rendim_2014_original)
```

################################################################################
### 2. Import Performance datasets

The databases are filtered keeping only cases with `MRUN` contained in the `Enrolment in Higher Education` databases. This filters out a total of 354,062 students enrolled in Higher education that are not present in the Performance databases. Approximately 95% of these filtered students have dates of birth before 1984, which is 18 years before the first Performance database. This implies that the students are not in the database because they had already finished their secondary education.

**DO NOT RUN THIS**
```{r, eval=FALSE}
# Import Performance dataset year by year
#-------------------------------------------------------------------------------
ImportRendimOne <- function(i) {
  # Vector of grades equivalent to 7th or 8th grade
  sept_octavo <- c('110-7', '110-8', '160-4', '160-5', '161-4', '161-5',
                   '163-4', '163-5', '165-3', '167-3')
  # create file path
  csv_filepath <- paste0("2_data/data_rendimiento/rendimiento_",i, "_publ.csv.gz")
  # read csv from compressed file
  temp_obj <- fread(csv_filepath, sep = ";", encoding = "UTF-8")
  # variable names to lowercase
  names(temp_obj) <- tolower(names(temp_obj))
  # remove cases not enroled in higher education (from ed.superior databases)
  temp_obj <- temp_obj[temp_obj$mrun %in% dropout_data$mrun]
  # Import only cases from 7th grade onward
  cod_ense_grado <- paste(temp_obj$cod_ense, temp_obj$cod_grado, sep = "-")
  temp_obj <- temp_obj[cod_ense_grado %in% sept_octavo | cod_ense > 300]
  # remove cases with situation "retirado" or "trasladado"
  temp_obj <- temp_obj[temp_obj$sit_fin_r == "P" | temp_obj$sit_fin_r == "R"]
  # order cases by "sit_fin" (so approved cases are first) and "asistencia"
  temp_obj <- temp_obj[order(temp_obj$sit_fin_r, -temp_obj$asistencia)]
  # remove duplicated mrun
  temp_obj <- temp_obj[!duplicated(temp_obj$mrun)]
  # change decimal separator from "," to "."
  temp_obj$prom_gral <- as.numeric(gsub(",", ".", temp_obj$prom_gral))
  return(temp_obj)
}
# Import data "rendimiento" one year
rendim_2019 <- ImportRendimOne(2019)
#head(rendim_2007, n=10)
rm(ImportRendimOne)
```

```{r, eval=FALSE}
# Import Performance databases from 2002 to 2019
#-------------------------------------------------------------------------------
ImportRendimAll <- function() {
  # Vector to select variables to keep
  vars_keep <- c("mrun", "rbd", "gen_alu", "fec_nac_alu", "cod_com_alu",
                 "cod_ense", "cod_grado", "prom_gral", "asistencia", "sit_fin")
  # Vector of grades equivalent to 7th or 8th grade
  sept_octavo <- c('110-7', '110-8', '160-4', '160-5', '161-4', '161-5',
                   '163-4', '163-5', '165-3', '167-3')
  # Loop to import and merge data
  for (i in 2002:2019) {
    # create file path
    csv_filepath <- paste0("2_data/data_rendimiento/rendimiento_",i, "_publ.csv.gz")
    # read csv from compressed file
    temp_obj <- fread(csv_filepath, sep = ";", encoding = "UTF-8")
    # variable names to lowercase
    names(temp_obj) <- tolower(names(temp_obj))
    # remove cases not enroled in higher education (from ed.superior databases)
    temp_obj <- temp_obj[temp_obj$mrun %in% dropout_data$mrun]
    # Import only cases from 7th grade onward
    cod_ense_grado <- paste(temp_obj$cod_ense, temp_obj$cod_grado, sep = "-")
    temp_obj <- temp_obj[cod_ense_grado %in% sept_octavo | cod_ense > 300]
    # remove cases with situation "retirado" or "trasladado"
    temp_obj <- temp_obj[temp_obj$sit_fin_r == "P" | temp_obj$sit_fin_r == "R"]
    # order cases by "sit_fin" (so approved cases are first) and "asistencia"
    temp_obj <- temp_obj[order(temp_obj$sit_fin_r, -temp_obj$asistencia)]
    # remove duplicated mrun
    temp_obj <- temp_obj[!duplicated(temp_obj$mrun)]
    # change decimal separator from "," to "."
    temp_obj$prom_gral <- as.numeric(gsub(",", ".", temp_obj$prom_gral))
    # select variables to keep
    temp_obj <- temp_obj[, ..vars_keep]
    # change names of variables (except mrun) to include year (e.g. "_2007")
    names(temp_obj) <- c("mrun", paste0(names(temp_obj[,2:length(temp_obj)]), "_", i))
    # merge all the datasets into one
    if (i == 2002) {
      rendim_all <- temp_obj
      }else {
        setkey(rendim_all, mrun)
        rendim_all <- merge(rendim_all, temp_obj, by = "mrun", all = TRUE)
      }
    }
  return(rendim_all)
  }
# Import data "rendimiento" 2002-2019
rendim_all <- ImportRendimAll()
rm(ImportRendimAll)
# Save Object
saveRDS(rendim_all, file = "2_data/saved_R_objects/03_rendim_2002_2019.RDS")
```

################################################################################
### 3. Merge databases

```{r}
# Load data and Merge into single DT
#-------------------------------------------------------------------------------
# Load dropout data and modify variables
dropout_data <- readRDS("2_data/saved_R_objects/02_dropout_data_2009_2018.RDS")
# Load performance data
rendim_all <- readRDS("2_data/saved_R_objects/03_rendim_2002_2019.RDS")
# Set keys
data.table::setkey(rendim_all, mrun)
data.table::setkey(dropout_data, mrun)
# Merge Dropout Data and Performance Data
drop_dt <- merge(rendim_all, dropout_data, by = "mrun", all.x = TRUE)
rm(dropout_data, rendim_all)
```

################################################################################
### 4. Make functions to obtain data

```{r}
# Make functions to obtain data
#-------------------------------------------------------------------------------
# function to get the year a student graduates from highschool
GetLastYear <- function(x) {
  max(c(2002:2019)[!is.na(x)])
  }
# function to get data from the year a student graduates from highschool
GetLastData <- function(x) {
  datalast = ifelse(all(is.na(x)), NA, x[max(which(!is.na(x)))])
  return(datalast)
  }
# function to get data from the first year with data
GetFirstData <- function(x) {
  datafirst = ifelse(all(is.na(x)), NA, x[min(which(!is.na(x)))])
  return(datafirst)
  }
# function to get the Average ignoring NA
GetAverage <- function(x){
  x = replace(x, which(x==0), NA)
  return(mean(x, na.rm = TRUE))
  }
# function to get the Median ignoring NA
GetMedian <- function(x){
  x = replace(x, which(x==0), NA)
  return(median(x, na.rm = TRUE))
  }
# function to get the number of unique data in period
GetNumUniq <- function(x) {
  length(unique(x[!is.na(x)]))
  }
# function to get the most common data in period
GetMostCommon <- function(x) {
  names(which.max(table(x[!is.na(x)])))
  }
# function to get the number of years failed in high school
GetNumFail <- function(x) {
  sum(x[!is.na(x)]=="R")
  }
# function to get the years a student is not enroled in school before finishing
GetNumOut <- function(x) {
  notna = !is.na(x)
  index_fst = min(c(1:18)[notna])
  index_lst = max(c(1:18)[notna])
  yrs_out = sum(is.na(x[index_fst:index_lst]))
  return(yrs_out)
  }
```

################################################################################
### 5. Get High-school Performance data for each student

```{r}
# Get highschool data for each student
#-------------------------------------------------------------------------------
# Making vector of RBD to subset columns
rbd_vec <- paste0("rbd_", 2002:2019)
# Add variable with the RBD of the last year for each student
drop_dt$hs_rbd <- apply(drop_dt[,..rbd_vec], 1, GetLastData)
# Add variable with the year each student graduates from highschool
drop_dt$hs_grad_year <- apply(drop_dt[,..rbd_vec], 1, GetLastYear)
# Making vector of "cod_ense" to subset columns
cod_ense_vec <- paste0("cod_ense_", 2002:2019)
# Adding variables with the last and first cod_ense for each student
drop_dt$hs_cod_ense <- apply(drop_dt[,..cod_ense_vec], 1, GetLastData)
drop_dt$hs_cod_ense_1st <- apply(drop_dt[,..cod_ense_vec], 1, GetFirstData)
# Making vector of "cod_grado" to subset columns
cod_grado_vec <- paste0("cod_grado_", 2002:2019)
# Adding variables with last and first cod_grado for each student
drop_dt$hs_cod_grado <- apply(drop_dt[,..cod_grado_vec], 1, GetLastData)
drop_dt$hs_cod_grado_1st <- apply(drop_dt[,..cod_grado_vec], 1, GetFirstData)
# Remove vectors
rm(rbd_vec, cod_ense_vec, cod_grado_vec)
```

```{r}
# Analysis of trajectory in highschool
#-------------------------------------------------------------------------------
# Table of years of graduation vs Year of enrolment in higher education
round(prop.table(table("HighEd"=drop_dt$year_1st_data,
                       "School"=drop_dt$hs_grad_year),1)*100,1)
```

```{r}
# Get performance data for each student
#-------------------------------------------------------------------------------
# Make vector of "prom_gral" to subset columns
prom_gral_vec <- paste0("prom_gral_", 2002:2019)
# Changing GPA 0 to NA
drop_dt[, (prom_gral_vec):= lapply(.SD, function(x) replace(x, which(x==0), NA)), .SDcols = prom_gral_vec]
# Add variable with the Average of the GPA of highschool for each student
drop_dt$hs_gpa_avg <- apply(drop_dt[,..prom_gral_vec], 1, GetAverage)
# Make vector of "asistencia" to subset columns
asistencia_vec <- paste0("asistencia_", 2002:2019)
# Determine number of cases with attendance > 100 for each year
#drop_dt[ , apply(.SD, 2, function(x) sum(x>100, na.rm = TRUE)), .SDcols = asistencia_vec]
# Fix 16 cases of students with attendance over 100
error_cases <- drop_dt$asistencia_2009>100 & !is.na(drop_dt$asistencia_2009)
error_attend <- drop_dt$asistencia_2009[error_cases]/10
drop_dt$asistencia_2009[error_cases] <- round(error_attend,0)
rm(error_attend, error_cases)
# Changing Attendance 0 to NA
drop_dt[, (asistencia_vec):= lapply(.SD, function(x) replace(x, which(x==0), NA)), .SDcols = asistencia_vec]
# Add variable with the Average attendance of highschool for each student
drop_dt$hs_attend_avg <- apply(drop_dt[,..asistencia_vec], 1, GetAverage)
# Add variable with the Median attendance of highschool for each student
drop_dt$hs_attend_med <- apply(drop_dt[,..asistencia_vec], 1, GetMedian)
# Make vector of "sit_fin" to subset columns
sit_fin_vec <- paste0("sit_fin_", 2002:2019)
# Add variable number of failed levels in highschool for each student
drop_dt$hs_fail_years <- apply(drop_dt[,..sit_fin_vec], 1, GetNumFail)
# Making vector of RBD to subset columns
rbd_vec <- paste0("rbd_", 2002:2019)
# Add variable of number of years out of the system before finishing
drop_dt$hs_years_out <- apply(drop_dt[,..rbd_vec], 1, GetNumOut)
# Remove vectors
rm(prom_gral_vec, asistencia_vec, sit_fin_vec, rbd_vec)
# Recoding education type variable
drop_dt <- drop_dt %>% mutate(
  hs_ed_type = case_when(
      hs_cod_ense == 310 ~ "Regular",
      hs_cod_ense == 410 ~ "Vocational",
      hs_cod_ense == 510 ~ "Vocational",
      hs_cod_ense == 610 ~ "Vocational",
      hs_cod_ense == 710 ~ "Vocational",
      hs_cod_ense == 810 ~ "Vocational",
      hs_cod_ense == 910 ~ "Vocational",
      hs_cod_ense < 400 ~ "Adult Regular",
      hs_cod_ense < 999 ~ "Adult Vocational"))
# Make factor from education type
drop_dt$hs_ed_type <- as.factor(drop_dt$hs_ed_type)
drop_dt$hs_ed_type <- relevel(drop_dt$hs_ed_type, ref = 3) # Reference: Regular
# Back to data.table
drop_dt <- data.table(drop_dt)
```

```{r}
# Determine how many years of data are available for each student
#-------------------------------------------------------------------------------
rbd_vec <- paste0("rbd_", 2002:2019)
drop_dt$num_yrs_data <- apply(drop_dt[,..rbd_vec], 1, function(x) sum(!is.na(x)))
message("Cases by number years with data in databases")
cbind(perc=round(prop.table(table(drop_dt$num_yrs_data))*100,1), num=table(drop_dt$num_yrs_data))
rm(rbd_vec)
```

################################################################################
### 6. Retrieve High School-level data

```{r}
# Retrieve data from databases High School 
#-------------------------------------------------------------------------------
ee_filepath <- paste0("2_data/data_establecimientos/directorio_oficial_ee_2019_20190430_publ.csv.gz")
# read csv from compressed file
ee_2019 <- fread(ee_filepath, sep = ";", encoding = "UTF-8")
# change variables to lower
names(ee_2019) <- tolower(names(ee_2019))
# recode variable of school control
ee_2019 <- ee_2019 %>%
    mutate(hs_control = case_when(
      cod_depe == 1 ~ "Public",
      cod_depe == 2 ~ "Public",
      cod_depe == 3 ~ "Charter",
      cod_depe == 4 ~ "Private",
      cod_depe == 5 ~ "CAD",
      cod_depe == 6 ~ "Public"))
# Make factor from hs_control
ee_2019$schl_control <- as.factor(ee_2019$hs_control)
# subset useful variables
ee_2019 <- ee_2019[, c("rbd", "cod_reg_rbd", "cod_com_rbd", "hs_control", "rural_rbd", "convenio_pie")]
# rename rbd to merge 
ee_2019 <- ee_2019 %>% rename(hs_rbd = rbd, hs_cod_reg = cod_reg_rbd, hs_cod_com = cod_com_rbd,
                              hs_rural = rural_rbd, hs_pie = convenio_pie)
# merge with "rendimiento" database
setkey(drop_dt, hs_rbd)
drop_dt <- merge(drop_dt, ee_2019, by = "hs_rbd", x.all = TRUE)
# Transform to factor High School-level variables
drop_dt$hs_control <- as.factor(drop_dt$hs_control)
drop_dt$hs_control <- relevel(drop_dt$hs_control, ref = 4) # Reference: Public
drop_dt$hs_rural <- as.factor(drop_dt$hs_rural)
drop_dt$hs_pie <- as.factor(drop_dt$hs_pie)
rm(ee_2019, ee_filepath)
```


################################################################################
### 7. Retrieve student-level data and check consistency across sources

```{r}
# Get data of students' gender from Performance Databases
#-------------------------------------------------------------------------------
# Create vector with columns names
gen_alu_vec <- paste0("gen_alu_", 2002:2019)
# Get gender in last year of highschool
stud_gender <- drop_dt[ , apply(.SD, 1, GetLastData), .SDcols = gen_alu_vec]
# Get number of genders in data for each student
gend_number <- drop_dt[ , apply(.SD, 1, GetNumUniq), .SDcols = gen_alu_vec]
# Get most common gender in data for students
stud_gend_max <- drop_dt[which(gend_number>1), apply(.SD, 1, GetMostCommon), .SDcols = gen_alu_vec]
stud_gender[gend_number>1] <- stud_gend_max
# Make new column for data with most common gender
drop_dt[, stud_gend2 := stud_gender]
drop_dt$stud_gend2[drop_dt$stud_gend2=="1"] <- "male"
drop_dt$stud_gend2[drop_dt$stud_gend2=="2"] <- "female"
drop_dt$stud_gend2 <- as.factor(drop_dt$stud_gend2)
rm(stud_gender, gend_number, stud_gend_max)
# Tables of proportion of inconsistent cases
message("Same gender in dropout data and performance data")
cbind("prop"=round(prop.table(table(drop_dt$stud_gend==drop_dt$stud_gend2, useNA = "ifany")),3),
      "num"=table(drop_dt$stud_gend==drop_dt$stud_gend2, useNA = "ifany"))
# Remove unnecesary columns
rm(gen_alu_vec)
```

```{r}
# Get  data of students' date of birth from Performance Databases
#-------------------------------------------------------------------------------
# Transform variables dob 2002 to same format as other years
drop_dt$fec_nac_alu_2002 <- as.Date(drop_dt$fec_nac_alu_2002, "%b %d %Y")
drop_dt$fec_nac_alu_2002 <- format(drop_dt$fec_nac_alu_2002,"%Y%m%d")
drop_dt$fec_nac_alu_2002 <- as.integer(drop_dt$fec_nac_alu_2002)
# Create vector with columns names
fec_nac_alu_vec <- paste0("fec_nac_alu_", 2002:2019)
# Get dob in last year of highschool
stud_dob <- drop_dt[ , apply(.SD, 1, GetLastData), .SDcols = fec_nac_alu_vec]
# Get number of dob in data for each student
dob_number <- drop_dt[ , apply(.SD, 1, GetNumUniq), .SDcols = fec_nac_alu_vec]
# Get most common dob in data for students
stud_dob_max <- drop_dt[which(dob_number>1), apply(.SD, 1, GetMostCommon), .SDcols = fec_nac_alu_vec]
stud_dob[dob_number>1] <- stud_dob_max
rm(stud_dob_max)
# Make new column for data with most common dob
drop_dt[, stud_date_birth2 := as.Date(stud_dob, "%Y%m%d")]
drop_dt[, stud_year_birth2 := as.numeric(format(drop_dt$stud_date_birth2, "%Y"))]
drop_dt[, stud_dob_consist_perf := dob_number]
rm(stud_dob, dob_number, fec_nac_alu_vec)
# Tables of proportion of inconsistent cases
message("Same year of birth in dropout data and performance data")
cbind("prop"=round(prop.table(table(drop_dt$stud_year_birth==drop_dt$stud_year_birth2, useNA = "ifany")),3),
      "num"=table(drop_dt$stud_year_birth==drop_dt$stud_year_birth2, useNA = "ifany"))
# Paste DOB data from Performance data to NA cases in Higher Ed. Data
drop_dt$stud_date_birth[is.na(drop_dt$stud_date_birth)] <- drop_dt$stud_date_birth2[is.na(drop_dt$stud_date_birth)]
drop_dt[, stud_year_birth := as.numeric(format(drop_dt$stud_date_birth, "%Y"))]

# Get  data based on students' date of birth
#-------------------------------------------------------------------------------
# Make variable of age at last year of high school
drop_dt$hs_grad_age <- drop_dt$hs_grad_year-drop_dt$stud_year_birth
# Make variable of age of enrolment in higher education
drop_dt$enrol_he_age <- drop_dt$year_1st_data-drop_dt$stud_year_birth
# Gap of years between graduation from highschool and enrolment in higher education
drop_dt$enrol_gap <- drop_dt$year_1st_data-drop_dt$hs_grad_year
# Table of distribution of data by source
yob_breaks <- c(1800,1900,1950,1960,1970,1980,1990,2000,2010)
yob_labels <- c("1800-1900","1900-1950","1950-1960","1960-1970","1970-1980",
                "1980-1990","1990-2000","2000-2010")
message("Year of birth by source of data")
cbind("Higher Ed. Data"=table(cut(drop_dt$stud_year_birth, yob_breaks, labels = yob_labels), useNA = "always"),
  "Performance Data"=table(cut(drop_dt$stud_year_birth2, yob_breaks, labels = yob_labels), useNA = "always"))
rm(yob_breaks, yob_labels)
```

```{r}
# Get  data of students' Comuna from Performance Databases
#-------------------------------------------------------------------------------
# Create vector with columns names
cod_com_alu_vec <- paste0("cod_com_alu_", 2002:2019)
# Get comuna from last year of highschool
stud_comuna <- drop_dt[ , apply(.SD, 1, GetLastData), .SDcols = cod_com_alu_vec]
# Get number of unique comunas in data for each student
comuna_num <- drop_dt[ , apply(.SD, 1, GetNumUniq), .SDcols = cod_com_alu_vec]
# Get most common comuna in data for students
stud_comuna_max <- drop_dt[which(comuna_num>1), apply(.SD, 1, GetMostCommon), .SDcols = cod_com_alu_vec]
stud_comuna[comuna_num>1] <- stud_comuna_max
# Make new column for data with most common comuna
drop_dt[, stud_comuna_num := comuna_num]
drop_dt[, stud_comuna := stud_comuna]
drop_dt$stud_comuna <- as.factor(drop_dt$stud_comuna)
rm(cod_com_alu_vec, stud_comuna, stud_comuna_max, comuna_num)
```

################################################################################
### 8. Change classes of variables

```{r}
setnames(drop_dt, old = "year_1st_data", new = "enrol_year")
drop_dt$inst_cruch <- factor(drop_dt$inst_cruch, levels = c(0, 1), labels = c("Not-cruch", "Cruch"))
drop_dt$hs_rural <- factor(drop_dt$hs_rural, levels = c(0, 1), labels = c("Urban", "Rural"))
drop_dt$hs_pie <- factor(drop_dt$hs_pie, levels = c(0, 1), labels = c("Not-PIE", "PIE"))
drop_dt$hs_rbd <- as.factor(drop_dt$hs_rbd)
drop_dt$hs_cod_ense <- as.factor(drop_dt$hs_cod_ense)
drop_dt$hs_cod_reg <- as.factor(drop_dt$hs_cod_reg)
drop_dt$hs_cod_com <- as.factor(drop_dt$hs_cod_com)
drop_dt$enrol_he_age <- as.integer(drop_dt$enrol_he_age)
drop_dt$stud_year_birth <- as.integer(drop_dt$stud_year_birth)
```

################################################################################
### 9. Save data

```{r}
# Set order and subset data
#-------------------------------------------------------------------------------
# rename program length and plan
setnames(drop_dt, old = "progr_len", new = "prog_len")
setnames(drop_dt, old = "progr_plan", new = "prog_plan")
# set order and subset
keep_vars <- c(
  # Students' characteristics
  "mrun", "stud_comuna",
  "stud_gend", "stud_gend2", "stud_date_birth", "stud_year_birth",
  "stud_date_birth2", "stud_year_birth2",
  # Quality control of student data
  "num_yrs_data", "stud_comuna_num", "stud_gend_consist", 
  "stud_dob_consist", "stud_dob_consist_perf",
  # Trajectory in Higher Education
  "enrol_year", "year_ori_enrol", "enrol_he_age", "enrol_gap", "num_uniq_prog",
  "dropout", "excluded", "change", "drop_change", "withdraw",
  # Higher Education Institution characteristics
  "inst_code", "inst_type", "inst_control", "inst_cruch", "inst_accred",
  # Higher Education Programme characteristics
  "prog_code", "prog_code_2nd", "prog_name", "prog_jornada", "prog_level",
  "prog_tuition", "prog_enrol_fee", "prog_accredit", "prog_plan", "prog_len",
  "prog_area1", "prog_area2", "prog_area3",
  # Trajectory in highschool
  "hs_grad_year", "hs_cod_ense", "hs_cod_grado", "hs_cod_ense_1st",
  "hs_cod_grado_1st",
  # High school characteristics
  "hs_rbd", "hs_ed_type", "hs_cod_reg", "hs_cod_com", "hs_control", "hs_rural",
  "hs_pie",
  # Students' Performance in highschool
  "hs_gpa_avg", "hs_attend_avg", "hs_attend_med", "hs_fail_years", "hs_years_out",
  "hs_grad_age"
  )
setcolorder(drop_dt, keep_vars)
drop_dt <- drop_dt[, keep_vars, with = FALSE]
rm(keep_vars)
```

```{r}
# Save Object
saveRDS(drop_dt, file = "2_data/saved_R_objects/04_dropout_rendim_data_2009_2018.RDS")
```
